---
title: "DESeq pair - compare"
author: "Ming Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float:
      collapsed: no
    keep_md: true
  word_document:
    toc: yes
  pdf_document:
    toc: yes
params:
  x: ""
  y: ""
  shrink_method: "apeglm"
---

```{r setup, include=FALSE}
suppressPackageStartupMessages(library(knitr))

knitr::opts_chunk$set(fig.width  = 12, 
                      fig.height = 8, 
                      fig.path   = "Figures/",
                      eval       = TRUE,
                      echo       = FALSE,
                      cache      = FALSE,
                      prompt     = FALSE,
                      tidy       = FALSE,
                      comment    = NA,
                      message    = FALSE,
                      warning    = FALSE,
                      eval       = T,
                      rownames.print = FALSE)
```


```{r load pkg}
library(DESeq2)
library(hiseqr)
library(knitr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(glue)
library(DT)
library(ggvenn)
library(RColorBrewer)
library(ggalluvial)
library(patchwork)
library(ggthemes)
```


```{r config}
# setwd("/data/yulab/wangming/work/yu_2020/projects/20200629_lxh_RNAseq/results/20200702_rnaseq_cmp/aaa")
# feature <- "gene"
# dirA    <- "../data/rnaseq_0310/shWhite_6h.vs.shAub-shAgo3_6h/"
# dirB    <- "../data/rnaseq_0310/shWhite_6h.vs.shPiwi-1_6h/"
x <- params$x
y <- params$y
shrink_method <- params$shrink_method

# x <- "/data/yulab/wangming/work/yu_2021/piwi_lxh/results/hiseq/RNAseq/Da/salmon/RNAseq_Da_shWhite_12h.vs.RNAseq_Da_shCG3893_12h"
# y <- "/data/yulab/wangming/work/yu_2021/piwi_lxh/results/hiseq/RNAseq/Da/salmon/RNAseq_Da_shWhite_12h.vs.RNAseq_Da_shPiwi_12h"
# shrink_method <- "apeglm"
## check
if(!all(is_hiseq_dir(c(x, y), "rnaseq_rx"))) {
  stop("not rnaseq_rx dir")
}
```


```{r}
pd_rds <- "Figures/deseq_pair.rds"
if(file.exists(pd_rds)) {
  pd <- readRDS(pd_rds)
} else {
  pd <- read_deseq_pair(x, y, shrink_method = shrink_method)
  if(! dir.exists(dirname(pd_rds))) {
    dir.create(dirname(pd_rds))
  }
  saveRDS(pd, pd_rds)
}
```

## Note

```{r, results='asis'}
msg <- glue::glue("Sample info",
                  "GroupA: {pd$x_smpname}",
                  "GroupB: {pd$y_smpname}",
                  "shrink_method: {shrink_method}")
print(msg)
```

## 1. Overall

```{r}
px <- stat_deseq_pair(x, y, .pd_rds = pd_rds, keep_not_sig = FALSE,
                      shrink_method = shrink_method, group = "all")
```

## 2.1 Number of significant changed genes

```{r}
hiseqr::to_DT(px$df1)
```

## 1.2 Overlap between groups

```{r, fig.width = 10, fig.height = 4}
p1 <- plot_overlap_deseq_pair(.pd_rds = pd_rds, group = "all")
p1$merge
```


## 2 Gene

```{r}
px <- stat_deseq_pair(x, y, .pd_rds = pd_rds, keep_not_sig = FALSE,
                      shrink_method = shrink_method, group = "gene")
```


## 2.1 Number of significant changed genes

```{r}
hiseqr::to_DT(px$df1)
```

## 2.2 Overlap between groups

```{r, fig.width = 10, fig.height = 4}
p1 <- plot_overlap_deseq_pair(.pd_rds = pd_rds, group = "gene")
p1$merge
```

## 2.3 Heatmap - sig genes

```{r, fig.width = 5, fig.height = 8}
plot_heatmap_deseq_pair(.pd_rds = pd_rds, group = "gene")
```

## 2.4 Alluvial - sig genes

```{r, fig.width = 6, fig.height = 4}
plot_alluvial_deseq_pair(.pd_rds = pd_rds, group = "gene")
```

## 2.5 Sig gene patterns

```{r}
df2 <- px$df3 %>%
  dplyr::group_by(trend) %>%
  dplyr::summarise(count = n())
hiseqr::to_DT(df2)
```

## 2.6 Sig gene list

```{r}
df3a <- px$pd$x_res %>%
  dplyr::select(gene_id, SYMBOL) %>%
  dplyr::mutate(across(everything(), as.character))
df3  <- dplyr::left_join(df3a, px$df2, by = "gene_id")
hiseqr::to_DT(df3, 1, pageLength = 20)
```


## 3 TE

```{r}
px <- stat_deseq_pair(x, y, .pd_rds = pd_rds, keep_not_sig = FALSE,
                      shrink_method = shrink_method, group = "te")
```

## 3.1 Number of significant changed genes

```{r}
hiseqr::to_DT(px$df1)
```

## 3.2 Overlap between groups

```{r, fig.width = 10, fig.height = 4}
p1 <- plot_overlap_deseq_pair(.pd_rds = pd_rds, group = "te")
p1$merge
```

## 3.3 Heatmap - sig genes

```{r, fig.width = 5, fig.height = 8}
plot_heatmap_deseq_pair(.pd_rds = pd_rds)
```

## 3.4 Alluvial - sig genes

```{r, fig.width = 6, fig.height = 4}
plot_alluvial_deseq_pair(.pd_rds = pd_rds, group = "te")
```

## 3.5 Sig gene patterns

```{r}
df2 <- px$df3 %>%
  dplyr::group_by(trend) %>%
  dplyr::summarise(count = n())
hiseqr::to_DT(df2)
```

## 3.6 Sig gene list

```{r}
df3a <-  px$pd$x_res %>%
  dplyr::select(gene_id, SYMBOL) %>%
  dplyr::mutate(across(everything(), as.character))
df3  <- dplyr::left_join(df3a, px$df2, by = "gene_id")
hiseqr::to_DT(df3, 1, pageLength = 20)
```


## 4 piRNA clusters

```{r}
px <- stat_deseq_pair(x, y, .pd_rds = pd_rds, keep_not_sig = FALSE,
                      shrink_method = shrink_method, group = "cluster")
```

## 4.1 Number of significant changed genes

```{r}
hiseqr::to_DT(px$df1)
```

## 4.2 Overlap between groups

```{r, fig.width = 10, fig.height = 4}
p1 <- plot_overlap_deseq_pair(.pd_rds = pd_rds, group = "cluster")
p1$merge
```

## 4.3 Heatmap - sig genes

```{r, fig.width = 5, fig.height = 8}
plot_heatmap_deseq_pair(.pd_rds = pd_rds)
```

## 4.4 Alluvial - sig genes

```{r, fig.width = 6, fig.height = 4}
plot_alluvial_deseq_pair(.pd_rds = pd_rds, group = "cluster")
```

## 4.5 Sig gene patterns

```{r}
df2 <- px$df3 %>%
  dplyr::group_by(trend) %>%
  dplyr::summarise(count = n())
hiseqr::to_DT(df2)
```

## 4.6 Sig gene list

```{r}
df3a <-  px$pd$x_res %>%
  dplyr::select(gene_id, SYMBOL) %>%
  dplyr::mutate(across(everything(), as.character))
df3  <- dplyr::left_join(df3a, px$df2, by = "gene_id")
hiseqr::to_DT(df3, 1, pageLength = 20)
```


**END**
