---
title: "Report: GO and KEGG enrichment analysis"
author: "Ming Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float:
      collapsed: no
    keep_md: true
  word_document:
    toc: yes
  pdf_document:
    toc: yes
params:
  input_dir: ""
---

```{r setup, include=FALSE}
suppressPackageStartupMessages(library(knitr))

knitr::opts_chunk$set(fig.width  = 12, 
                      fig.height = 8, 
                      fig.path   = "Figures/",
                      eval       = TRUE,
                      echo       = FALSE,
                      cache      = FALSE,
                      prompt     = FALSE,
                      tidy       = FALSE,
                      comment    = NA,
                      message    = FALSE,
                      warning    = FALSE,
                      eval       = T,
                      rownames.print = FALSE)
```


```{r load pkg}
library(hiseqr)
library(knitr)
library(ggplot2)
library(dplyr)
library(glue)
library(DT)
library(plotly)
library(reticulate)
library(ggvenn)
library(RColorBrewer)
library(ggalluvial)
library(patchwork)
```


```{r config}
input_dir <- normalizePath(params$input_dir)
```


```{r}
# global vars
stopifnot(hiseqr::is_enrich_dir(input_dir))
pd <- hiseqr::read_enrich(input_dir)
```


## 1 Summary

```{r}      
s <- glue::glue("GO analysis for [ {names(pd)} ] gene set")
print(s)
```


### 1.1 Number of genes

Here are the numbers of significant changed genes

```{r, eval = T}
go_data <- pd[[1]]$go_data.rds
pd2 <- readRDS(go_data)
gene_list <- pd2$gene_list
msg <- glue::glue("Input genes: {length(gene_list)}")
print(msg)
```

## 2. GO analysis

```{r}
# data
f1 <- pd[[1]]$go_plots.rds
pdata <- readRDS(f1)
```

### 2.1 Overview of gene **Categories**

```{r, fig.width = 10, fig.height = 6}
pdata$group$p_wego
```

### 2.2 Overview of **Enriched** genes

```{r, fig.width = 10, fig.height = 6}
pdata$enrich$p_wego
```

### 2.3 Gene **Categories**

```{r, fig.width = 8, fig.height = 10, eval = T}
# barplot
p1 <- pdata$group$BP$barplot
p2 <- pdata$group$CC$barplot
p3 <- pdata$group$MF$barplot

# p1 / p2 / p3
cowplot::plot_grid(p1, p2, p3, ncol = 1)
```


### 2.4 Enriched genes - dotplot

```{r, fig.width = 8, fig.height = 10}
# dotplot
p1 <- pdata$enrich$BP$dotplot
p2 <- pdata$enrich$CC$dotplot
p3 <- pdata$enrich$MF$dotplot

# p1 / p2 / p3
cowplot::plot_grid(p1, p2, p3, ncol = 1)
```


### 2.5 Enriched genes - cnetplot

```{r, fig.width = 8, fig.height = 10}
# dotplot
p1 <- pdata$enrich$BP$cnetplot
p2 <- pdata$enrich$CC$cnetplot
p3 <- pdata$enrich$MF$cnetplot

# p1 / p2 / p3
cowplot::plot_grid(p1, p2, p3, ncol = 1)
```


### 2.6 Enriched genes - emapplot

```{r, fig.width = 8, fig.height = 10, eval = FALSE, eval = FALSE}
# dotplot
p1 <- pdata$enrich$BP$emapplot  
p2 <- pdata$enrich$CC$emapplot
p3 <- pdata$enrich$MF$emapplot

# p1 / p2 / p3
cowplot::plot_grid(p1, p2, p3, ncol = 1)
```


## 3 KEGG analysis

```{r}
# data
f1 <- pd[[1]]$kegg_plots.rds
pdata <- readRDS(f1)
```

### 3.1 Overview of gene **Categories**

```{r, fig.width = 10, fig.height = 5}
pdata$enrich$p_wego
```


### 3.2 Gene **Categories**

```{r, fig.width = 6, fig.height = 5, eval = FALSE}
# barplot
p1 <- pdata$enrich$kegg$barplot
p1
```


### 3.3 Enriched genes - dotplot

```{r, fig.width = 6, fig.height = 5}
# dotplot
p1 <- pdata$enrich$kegg$dotplot
p1
```


### 3.4 Enriched genes - cnetplot

```{r, fig.width = 6, fig.height = 6}
# cnetplot
p1 <- pdata$enrich$kegg$cnetplot
p1
```


### 3.5 Enriched genes - emapplot

```{r, fig.width = 8, fig.height = 10, eval = FALSE, eval = FALSE}
# dotplot
p1 <- pdata$enrich$kegg$emapplot
p1
```

**END**
