---
title: "HiSeq Report"
author: "Ming Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float:
      collapsed: no
    keep_md: false
  word_document:
    toc: yes
  pdf_document:
    toc: yes
params:
  input_dir: ""
---

```{r setup, include=FALSE}
suppressPackageStartupMessages(library(knitr))

knitr::opts_chunk$set(fig.width  = 12, 
                      fig.height = 8, 
                      fig.path   = "Figures/",
                      eval       = TRUE,
                      echo       = FALSE,
                      cache      = FALSE,
                      prompt     = FALSE,
                      tidy       = FALSE,
                      comment    = NA,
                      message    = FALSE,
                      warning    = FALSE,
                      eval       = T,
                      rownames.print = FALSE)
```


```{r}
library(hiseqr)
library(DESeq2)
library(ggplot2)
library(patchwork)
library(fishualize)
library(dplyr)
library(glue)
library(patchwork)
```


```{r init, eval = T}
input_dir <- params$input_dir
# input_dir <- "../"
input_dir <- normalizePath(input_dir)
if(! is_hiseq_dir(input_dir, "rnaseq_rx")) {
  stop(paste0("Not a RNAseq directory, rnaseq_rx: ", input_dir))
}
```

## Config data

```{r config}
# check paired
is_paired <- all(list_hiseq_file(input_dir, "is_paired", "r1"))
n_samples <- length(list_hiseq_file(input_dir, "smp_name", "_r1"))
genome    <- list_hiseq_file(input_dir, "genome", "_rx")
organism  <- genome
gtf       <- list_hiseq_file(input_dir, "gene_gtf", "_rx")
wt_name   <- list_hiseq_file(input_dir, "wt_name", "_rx")
mut_name  <- list_hiseq_file(input_dir, "mut_name", "_rx")
project_name <- list_hiseq_file(input_dir, "project_name", "_rx")
aligner   <- list_hiseq_file(input_dir, "aligner")
is_star   <- aligner %in% c("STAR", "hisat2")

msg <- glue::glue(
  "Configuration: ",
  "   wildtype: {wt_name} ",
  "     mutant: {mut_name} ",
  "     genome: {genome} ",
  " paired-end: {is_paired} ",
  "    aligner: {aligner} ",
  "project_dir: {project_name}",
  .sep = "\n"
)
print(msg)
```

```{r plot_data, eval = is_star}
df_trim  <- read_hiseq_trim_stat(input_dir)
df_align <- read_hiseq_align_stat(input_dir)
```


```{r loading data, eval = is_star}
pd <- hiseq_summary(input_dir, TRUE)

# summary
n_smp      <- length(list_hiseq_file(input_dir, "rep_list", "auto"))
if(inherits(pd$align, "data.frame") & inherits(pd$align, "data.frame")) {
  total      <- df_align %>% pull(total) %>% sum()
  map        <- df_align %>% pull(map) %>% sum()
  map_pct    <- round(map / total * 100, 2) %>% sum()
  unique_m   <- df_align %>% pull(unique) %>% sum()
  unique_pct <- round(unique_m / total * 100, 1)
  total      <- round(total / 1e6, 2)
  map        <- round(map / 1e6, 2)
  unique_m   <- round(unique_m / 1e6, 1) # Million
} else {
  total <- map <- unique_m <- 1
  unique_pct <- map_pct <- 100
}
```

## 1 Summary

```{r summary, results="asis", eval = is_star}
sum_text <- glue::glue(
  "A total of {n_samples} samples; {total}M sequencing depth; ",
  "with {map}M ({map_pct}%) reads mapped to reference genome, including ",
  "{unique_m}M ({unique_pct}%) unique mapped."
)
print(sum_text)
```

## 2 Results


### 2.1 Table 1. Trimming reads

```{r, eval = is_star}
if(inherits(pd$trim, "data.frame")) {
  pd$trim %>%
    dplyr::select(name, input, output, out_pct, rm_pct) %>%
    dplyr::rename(clean = output, 
                  clean_pct = out_pct,
                  short_pct = rm_pct) %>%
    hiseqr::to_DT()
}
```

```{r figure1_trim_reads, fig.height=3, fig.width=6, eval = is_star}
if(inherits(pd$trim, "data.frame")) {
  plot_hiseq_trim(pd$trim)
}
```


### 2.2 Table 2. Mapping reads

```{r, eval = is_star}
if(inherits(pd$align, "data.frame")) {
  pd$align %>%
    dplyr::select(name, total, map, unmap) %>%
    dplyr::mutate(map_pct = round(map / total * 100, 2)) %>%
    hiseqr::to_DT()
}
```

```{r figure1_align_reads, fig.height=3, fig.width=6, eval = is_star}
if(inherits(pd$align, "data.frame")) {
  plot_hiseq_align(pd$align, columns = c("map", "rRNA", "unmap"))
}
```

+ unique and multiple mapping reads

```{r, eval = is_star}
if(inherits(pd$align, "data.frame")) {
pd$align %>%
  dplyr::select(name, total, unique, multi, unmap) %>%
  dplyr::mutate(unique_pct = round(unique / total * 100, 2)) %>%
  hiseqr::to_DT()
}
```

```{r unique_multi_reads, fig.height=3, fig.width=6, eval = is_star}
if(inherits(pd$align, "data.frame")) {
plot_hiseq_align(pd$align, columns = c("unique", "multi", "unmap"))
}
```


### 2.3 RNAseq library type

Check the RNAseq library type, check if the reads from forward strand or reverse strand of mRNA;

Using `featureCounts` to count reads on mRNAs.

For strand-specific library

| sens | anti | library | example | 
| -- | -- | -- | -- |
| -s 1 | -s 2 | `read1: ++ -- / read2: -- ++` | - | 
| -s 2 | -s 1 | `read1: +- -+ / read2: +- -+` | dUTP library, NSR library |


```{r lib type, results="asis", eval = is_star}
lib1 = "read1: ++ -- / read2: -- ++"
lib2 = "read1: +- -+ / read2: +- -+"
lib3 = "read1, read2, non-strand-specific"

f <- list_hiseq_file(input_dir, "strandness_json", "rnaseq_rn")
if(length(f) > 0) {
  s <- jsonlite::read_json(f[1])
  a <- switch(as.character(s$sense),
              "0" = {print(lib3)},
              "1" = {print(lib1)},
              "2" = {print(lib2)})
}
# print(a)
```

### 2.4 Quantification

```{r, results="asis", eval = is_star}
msg <- tryCatch(
  {
    glue::glue(
      "Count reads on features:",
      "Genome: {genome}",
      "   GTF: {basename(gtf)}",
      .sep = "\n"
    )
  },
  error=function(cond) {
    message("Failed reading quantification")
  }
)
print(msg)
```
**na**, indicate the `Not Assigned reads`.

```{r table2_sens_summary, eval = is_star}
p4 <- plot_rnaseq_featureCounts(input_dir)
if(is(p4, "ggplot")) {
  # show table
  df4 <- p4$data %>%
    dplyr::select(sample, total, strand, count) %>%
    tidyr::pivot_wider(names_from = strand, values_from = count) %>%
    dplyr::mutate(sens_pct = round(sense / total * 100, 1))
  hiseqr::to_DT(df4)
}
```


**na**, indicate the `Not Assigned reads`.
```{r count strand, fig.height=3, fig.width=6, eval = is_star}
print(p4)
```


### 2.5 Correlation between replicates

#### 1 Distribution of reads on gene body

#### 2 Correlation - PCA

#### 3 Correlation - heatmap

```{r, eval = is_star}
g1 <- list_hiseq_file(input_dir, "genebody_enrich_png", "_rx")
g2 <- list_hiseq_file(input_dir, "bam_cor_pca_png", "_rn")
g3 <- list_hiseq_file(input_dir, "bam_cor_heatmap_png", "_rn")
g  <- purrr::keep(c(g1, g2, g3), file.exists)
# if(file.exists(g)) {
if(length(g) > 0) {
  include_graphics(g)
}
```



## 3 Differentially expressed genes

### 3.1 Info

```{r}
msg <- glue::glue(
  "Configuration:",
  " Compare: log2( {mut_name} / {wt_name} )",
  " Project: {project_name}",
  "  Genome: {genome}",
  "wildtype: {wt_name}",
  "  mutant: {mut_name}",
  "quantification: featureCounts (v2.0.0)",
  "DE analysis: DESeq2 (v1.30.1)",
  "cutoff: padj < 0.05 and |log2fc| >= 1",
  .sep = "\n")
print(msg)
```

```{r read_deseq, results="asis"}
# data.frame
deseq_dir <- list_hiseq_file(input_dir, "deseq_dir", "_rx")
dds_trans <- deseq_qc_dds(deseq_dir, return_data = "dds_trans")
res_lfc   <- deseq_qc_dds(deseq_dir, return_data = "res_lfc")

# all data.frame
df_list   <- sapply(names(res_lfc), function(i) {
  deseq_qc_res(deseq_dir, shrink_method = i)
}, simplify = FALSE, USE.NAMES = TRUE)


get_deseq_sig_table <- function(x, ...) {
  if(!inherits(x, "list")) {
    return(NULL)
  }
  #for(name in names(x)) {
  lapply(names(x), function(name) {
    df <- x[[name]]
    if(inherits(df, "data.frame")) {
      if("sig" %in% names(df)) {
        as.data.frame(table(df[["sig"]])) %>%
          dplyr::mutate(sample = name)
      }
    }
  })%>%
    dplyr::bind_rows()
}

# function-3: get sig plot, by rnaseq_rx path
#' @param x list of data.frame, res table
get_deseq_sig_plot <- function(x, ...) {
  df <- get_deseq_sig_table(x, ...)
  if(inherits(df, "data.frame")) {
    df %>%
      dplyr::rename(sig = Var1) %>%
      dplyr::filter(!sig == "not") %>%
      ggplot(aes(sig, Freq, group = sample, fill = sample, label = Freq)) +
      geom_col(position = position_dodge()) +
      geom_text(position = position_dodge(.9), vjust = -.5) +
      fishualize::scale_fill_fish(discrete = TRUE, option = "Trimma_lantana") +
      xlab(NULL) +
      ylab("Number of genes") +
      ggtitle("Number of DE genes") +
      ggplot2::theme_bw()
  }
}


# # sig table
# df_sig <- lapply(names(df_list), function(i) {
#   as.data.frame(table(df_list[[i]][["sig"]])) %>%
#     dplyr::mutate(shrink = i)
# }) %>%
#   dplyr::bind_rows()
# df_deseq <- read_rnaseq_deseq2(input_dir)
# # add link to database: ENSEMBL, FlyBase
# df_deseq$link <- gene_to_link(df_deseq$Gene, organism, style = "html",
#                               site = "http://www.flybase.org", readable = TRUE)
```


**sample names table**

```{r}
smp_csv <- list_hiseq_file(deseq_dir, "smp_name_csv", "auto")
if(file.exists(smp_csv)) {
  tb <- read.csv(smp_csv)
} else {
  tb <- NULL
}
hiseqr::to_DT(tb)
```

### 3.2 Number of differentially expressed genes

Here are the numbers of significant changed genes, criteria: `foldChange >= 2, pvalue < 0.05`

```{r parsing de table}
get_deseq_sig_table(df_list) %>%
  filter(sample == "apeglm") %>%
  hiseqr::to_DT()
```

```{r, fig.width = 4, fig.height = 4}
get_deseq_sig_plot(df_list['apeglm'])
```


### 3.3 Up-regulated genes

```{r}
sig_gene_list <- function(sig_type = "up") {
  .col_select <- c("shrink", "gene_id")
  if("SYMBOL" %in% names(df_list[[1]])) {
    .col_select <- c(.col_select, "SYMBOL")
  }
  lapply(names(res_lfc), function(i) {
    df_list[[i]] %>%
      dplyr::filter(sig == sig_type) %>%
      dplyr::mutate(shrink = i) %>%
      dplyr::select(all_of(.col_select), 2:3, log2FoldChange, padj, sig) %>%
      # dplyr::select(shrink, 1:3, log2FoldChange, padj, sig) %>%
      dplyr::mutate(across(where(is.numeric), ~round(.x, digits = 1)))
  }) %>%
    dplyr::bind_rows()
}

## shrink
sig_gene_list("up") %>%
  filter(shrink == "apeglm") %>%
  hiseqr::to_DT(4)
```


### 3.4 Down-regulated genes

```{r}
## shrink
sig_gene_list("down") %>%
  filter(shrink == "apeglm") %>%
  hiseqr::to_DT(4)
```


## 4 DESeq quality control

### 4.1 PCA between samples

```{r, fig.width = 10, fig.height = 4}
p_list <- lapply(names(dds_trans), function(i) {
  deseq_qc_pca(deseq_dir, outdir = deseq_dir, transform_method = i)
})
patchwork::wrap_plots(p_list, nrow = 1, guides = "collect")
```

### 4.2 Sample distance

```{r, fig.width = 10, fig.height = 4}
p_list <- lapply(names(dds_trans), function(i) {
  deseq_qc_dist(deseq_dir, outdir = deseq_dir, transform_method = i)
})
patchwork::wrap_plots(p_list, nrow = 1, guides = "collect")
```

### 4.3 Mean standard deviation of baseMean

```{r, fig.width = 10, fig.height = 4}
p_list <- lapply(names(dds_trans), function(i) {
  deseq_qc_mean_sd(deseq_dir, outdir = deseq_dir, transform_method = i)
})
patchwork::wrap_plots(p_list, nrow = 1, guides = "collect")
```

### 4.4 MA plot

```{r, fig.width = 8, fig.height = 4}
# p_list <- lapply(names(res_lfc), function(i) {
p_list <- lapply("apeglm", function(i) {
  p <- deseq_qc_ma(deseq_dir, outdir = deseq_dir, shrink_method = i) +
    ggtitle(glue::glue("shrink: {i}"))
  deseq_qc_add_sig_label(p)
})
patchwork::wrap_plots(p_list, ncol = 2, guides = "collect")
```

### 4.5 Volcano plot

```{r, fig.width = 8, fig.height = 4}
# p_list <- lapply(names(res_lfc), function(i) {
p_list <- lapply("apeglm", function(i) {
  p <- deseq_qc_volcano(deseq_dir, outdir = deseq_dir, shrink_method = i) +
    ggtitle(glue::glue("shrink: {i}"))
  deseq_qc_add_sig_label(p)
})
patchwork::wrap_plots(p_list, ncol = 2, guides = "collect")
```

### 4.6 Scatter plot

```{r, fig.width = 8, fig.height = 4}
# p_list <- lapply(names(res_lfc), function(i) {
p_list <- lapply("apeglm", function(i) {
  p <- deseq_qc_scatter(deseq_dir, outdir = deseq_dir, shrink_method = i) +
    ggtitle(glue::glue("shrink: {i}"))
  deseq_qc_add_sig_label(p)
})
patchwork::wrap_plots(p_list, ncol = 2, guides = "collect")
```


## 5 Detail report

```{r, report_list}
data_dir <- "data"
check_path(data_dir)
# copy files
dirs <- list_hiseq_file(input_dir, "report_dir", hiseq_type = TRUE)
report_list <- c()
for(i in dirs) {
  src_html <- list_hiseq_file(dirname(i), "report_html", hiseq_type = "auto")
  smp_name <- list_hiseq_file(dirname(i), "smp_name", hiseq_type = "auto")
  dest_html <- file.path(data_dir, paste0(smp_name, ".report.html"))
  file.copy(src_html, dest_html, overwrite = TRUE, copy.date = TRUE)
  m <- list(html = dest_html)
  names(m) <- smp_name
  report_list <- c(report_list, m)
}

# to data.frame
if(length(report_list) > 0) {
  df_report <- data.frame(
    name = names(report_list),
    file = unlist(report_list)
  ) %>%
    dplyr::mutate(
      link = paste0("<a href=", file, " target='_blank'>", name, "</a>")) %>%
  dplyr::select(link) %>%
    tibble::remove_rownames()
  # to table
  hiseqr::to_DT(df_report, mode = 3)
}
```


**END**
