---
title: "HiSeq Report"
author: "Ming Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float:
      collapsed: no
    keep_md: false
  word_document:
    toc: yes
  pdf_document:
    toc: yes
params:
  input_dir: ""
---

```{r setup, include=FALSE}
suppressPackageStartupMessages(library(knitr))

knitr::opts_chunk$set(fig.width  = 12, 
                      fig.height = 8, 
                      fig.path   = "Figures/",
                      eval       = TRUE,
                      echo       = FALSE,
                      cache      = FALSE,
                      prompt     = FALSE,
                      tidy       = FALSE,
                      comment    = NA,
                      message    = FALSE,
                      warning    = FALSE,
                      eval       = T,
                      rownames.print = FALSE)
```


```{r}
library(hiseqr)
library(DESeq2)
library(ggplot2)
library(patchwork)
library(fishualize)
library(dplyr)
library(glue)
library(patchwork)
```


```{r init, eval = T}
input_dir  <- params$input_dir
# input_dir <- "~/work/devel_pipeline/hiseq/rnaseq/output/dm6_te_piRC/pe_control.vs.pe_treatment/"
input_dir  <- normalizePath(input_dir)
if(! is_hiseq_dir(input_dir, "rnaseq_rx")) {
  stop(paste0("Not a RNAseq directory, rnaseq_rx: ", input_dir))
}
```

## Config data

```{r config, eval = T, results = "asis"}
# check paired
is_paired <- all(list_hiseq_file(input_dir, "is_paired", "r1"))
n_samples <- length(list_hiseq_file(input_dir, "smp_name", "_r1"))
genome    <- list_hiseq_file(input_dir, "genome", "_rx")
organism  <- genome
gtf       <- list_hiseq_file(input_dir, "gene_gtf", "_rx")
wt_name   <- list_hiseq_file(input_dir, "wt_name", "_rx")
mut_name  <- list_hiseq_file(input_dir, "mut_name", "_rx")
project_name <- list_hiseq_file(input_dir, "project_name", "_rx")

msg <- glue::glue(
  "Configuration:",
  "   wildtype: {wt_name}",
  "     mutant: {mut_name}",
  "     genome: {genome}",
  " paired-end: {is_paired}",
  "project_dir: {project_name}",
  .sep = "\n"
)
print(msg)
```


```{r plot_data}
df_trim_stat  <- read_hiseq_trim_stat(input_dir)
df_align_stat <- read_hiseq_align_stat(input_dir)
```

## 1 Summary

```{r summary, results="asis"}
if(is(df_align_stat, "data.frame")) {
  total      <- df_align_stat %>% pull(total) %>% mean()
  map        <- df_align_stat %>% pull(map) %>% mean()
  map_pct    <- round(map / total * 100, 1) %>% mean()
  unique_m   <- df_align_stat %>% pull(unique) %>% mean()
  unique_pct <- round(unique_m / total * 100, 1) %>% mean()
  total    <- round(total / 1e6, 1) # Million
  map      <- round(map / 1e6, 1) # Million
  unique_m <- round(unique_m / 1e6, 1) # Million
  sum_text <- glue::glue("A total of {n_samples} samples; {total}M sequencing depth; ",
                         "with {map}M ({map_pct}%) reads mapped to reference genome, including ",
                         "{unique_m}M ({unique_pct}%) unique mapped.")
  
  print(sum_text)
}
```

## 2 Results


### 2.1 Table1. number of reads

```{r table1_mito_pct}
if(is(df_align_stat, "data.frame")) {
  hiseqr::to_DT(df_align_stat)
}
```


### 2.2 Figure 1. Number of mapped reads

+ Clean data percentage

```{r figure1_trim_reads, out.width = "100%", fig.height = 3}
p2_2 <- tryCatch(
  {
    plot_hiseq_trim(input_dir)
  },
  error=function(cond) {
    message("Failed plotting: 2.2.number_of_mapped_reads")
  }
)
print(p2_2)
```

+ Map percentage

```{r figure1_align_reads, fig.height = 3}
p2_3 <- tryCatch(
  {
    plot_hiseq_align(input_dir, columns = c("map", "rRNA", "spikein", "unmap"))
  },
  error=function(cond) {
    message("Failed plotting: 2.3.map_percentage")
  }
)
print(p2_3)
```

+ Unique percentage

```{r figure1_unique_reads, fig.height = 3}
p2_4 <- tryCatch(
  {
    plot_hiseq_align(input_dir, columns = c("unique", "multi", "unmap"))
  },
  error=function(cond) {
    message("Failed plotting: 2.4.unique_map")
  }
)
print(p2_4)
```


### 2.3 RNAseq library type

Check the RNAseq library type, check if the reads from forward strand or reverse strand of mRNA;

Using `featureCounts` to count reads on mRNAs.

For strand-specific library

| sens | anti | library | example | 
| -- | -- | -- | -- |
| -s 1 | -s 2 | `read1: ++ -- / read2: -- ++` | - | 
| -s 2 | -s 1 | `read1: +- ++ / read2: +- -+` | dUTP library, NSR library |


```{r lib type, results="asis"}
lib1 = "read1: ++ -- / read2: -- ++"
lib2 = "read1: +- ++ / read2: +- -+"
lib3 = "read1, read2, non-strand-specific"

f <- list_hiseq_file(input_dir, "strandness_json", "rnaseq_rn")
if(length(f) > 0) {
  s <- jsonlite::read_json(f[1])
  a <- switch(as.character(s$sense),
              "0" = {print(lib3)},
              "1" = {print(lib1)},
              "2" = {print(lib2)})
}
# print(a)
```

### 2.4 Quantification

```{r, results="asis"}
msg <- tryCatch(
  {
    glue::glue(
      "Count reads on features:",
      "Genome: {genome}",
      "   GTF: {basename(gtf)}",
      .sep = "\n"
    )
  },
  error=function(cond) {
    message("Failed reading quantification")
  }
)
print(msg)
```
**na**, indicate the `Not Assigned reads`.

```{r table2_sens_summary}
p4 <- plot_rnaseq_featureCounts(input_dir)
if(is(p4, "ggplot")) {
  # show table
  df4 <- p4$data %>%
    dplyr::select(sample, total, strand, count) %>%
    tidyr::pivot_wider(names_from = strand, values_from = count) %>%
    dplyr::mutate(sens_pct = round(sense / total * 100, 1))
  hiseqr::to_DT(df4)
}
```


**na**, indicate the `Not Assigned reads`.
```{r count strand, fig.height=3}
print(p4)
```


### 2.5 Correlation between replicates

#### 1 Distribution of reads on gene body

#### 2 Correlation - PCA

#### 3 Correlation - heatmap

```{r}
g1 <- list_hiseq_file(input_dir, "genebody_enrich_png", "_rx")
g2 <- list_hiseq_file(input_dir, "bam_cor_pca_png", "_rn")
g3 <- list_hiseq_file(input_dir, "bam_cor_heatmap_png", "_rn")
g  <- purrr::keep(c(g1, g2, g3), file.exists)
# if(file.exists(g)) {
if(length(g) > 0) {
  include_graphics(g)
}
```



## 3 Differentially expressed genes

### 3.1 Info

```{r , results="asis"}
msg <- glue::glue(
  "Configuration:",
  " Compare: log2( {mut_name} / {wt_name} )",
  " Project: {project_name}",
  "  Genome: {genome}",
  "wildtype: {wt_name}",
  "  mutant: {mut_name}",
  "quantification: featureCounts (v2.0.0)",
  "DE analysis: DESeq2 (v1.30.1)",
  "cutoff: pajd < 0.05 and |log2fc| >= 1",
  .sep = "\n")
print(msg)
```

```{r read_deseq, results="asis"}
# data.frame
deseq_dir <- list_hiseq_file(input_dir, "deseq_dir", "_rx")
dds_trans <- deseq_qc_dds(deseq_dir, return_data = "dds_trans")
res_lfc   <- deseq_qc_dds(deseq_dir, return_data = "res_lfc")

# all data.frame
df_list   <- sapply(names(res_lfc), function(i) {
  deseq_qc_res(deseq_dir, shrink_method = i)
  # read_rnaseq_deseq2(input_dir, shrink_method = i)
}, simplify = FALSE, USE.NAMES = TRUE)

# sig table
df_sig <- lapply(names(df_list), function(i) {
  as.data.frame(table(df_list[[i]][["sig"]])) %>%
    dplyr::mutate(shrink = i)
}) %>%
  dplyr::bind_rows()
# df_deseq <- read_rnaseq_deseq2(input_dir)
# # add link to database: ENSEMBL, FlyBase
# df_deseq$link <- gene_to_link(df_deseq$Gene, organism, style = "html",
#                               site = "http://www.flybase.org", readable = TRUE)
```


**sample names table**

```{r}
smp_csv <- list_hiseq_file(deseq_dir, "smp_name_csv", T)
if(file.exists(smp_csv)) {
  tb <- read.csv(smp_csv)
} else {
  tb <- NULL
}
hiseqr::to_DT(tb)
```

### 3.2 Number of differentially expressed genes

Here are the numbers of significant changed genes, criteria: `foldChange >= 2, pvalue < 0.05`

```{r parsing de table}
df_sig %>%
  tidyr::pivot_wider(names_from = "Var1", values_from = "Freq") %>%
  hiseqr::to_DT()
```

```{r, fig.width = 7, fig.height = 5}
df_sig %>%
  dplyr::filter(!Var1 == "not") %>%
  ggplot(aes(Var1, Freq, group = shrink, fill = shrink, label = Freq)) + 
  geom_col(position = position_dodge()) + 
  geom_text(position = position_dodge(.9), vjust = -.5) +
  fishualize::scale_fill_fish(discrete = TRUE, option = "Trimma_lantana") +
  xlab(NULL) +
  ylab("Number of genes") +
  ggtitle("Number of DE genes") +
  ggplot2::theme_bw()
```


### 3.3 Up-regulated genes

```{r}
sig_gene_list <- function(sig_type = "up") {
  .col_select <- c("shrink", "gene_id")
  if("SYMBOL" %in% names(df_list[[1]])) {
    .col_select <- c(.col_select, "SYMBOL")
  }
  lapply(names(res_lfc), function(i) {
    df_list[[i]] %>%
      dplyr::filter(sig == sig_type) %>%
      dplyr::mutate(shrink = i) %>%
      dplyr::select(all_of(.col_select), 2:3, log2FoldChange, padj, sig) %>%
      # dplyr::select(shrink, 1:3, log2FoldChange, padj, sig) %>%
      dplyr::mutate(across(where(is.numeric), ~round(.x, digits = 1)))
  }) %>%
    dplyr::bind_rows()
}

## shrink
hiseqr::to_DT(sig_gene_list("up"), 4)
```


### 3.4 Down-regulated genes

```{r}
hiseqr::to_DT(sig_gene_list("down"), 4)
```


## 4 DESeq quality control

### 4.1 PCA between samples

```{r, fig.width = 10, fig.height = 4}
p_list <- lapply(names(dds_trans), function(i) {
  deseq_qc_pca(deseq_dir, outdir = deseq_dir, transform_method = i)
})
patchwork::wrap_plots(p_list, nrow = 1, guides = "collect")
```

### 4.2 Sample distance

```{r, fig.width = 10, fig.height = 4}
p_list <- lapply(names(dds_trans), function(i) {
  deseq_qc_dist(deseq_dir, outdir = deseq_dir, transform_method = i)
})
patchwork::wrap_plots(p_list, nrow = 1, guides = "collect")
```

### 4.3 Mean standard deviation of baseMean

```{r, fig.width = 10, fig.height = 4}
p_list <- lapply(names(dds_trans), function(i) {
  deseq_qc_mean_sd(deseq_dir, outdir = deseq_dir, transform_method = i)
})
patchwork::wrap_plots(p_list, nrow = 1, guides = "collect")
```

### 4.4 MA plot

```{r, fig.width = 7, fig.height = 7}
p_list <- lapply(names(res_lfc), function(i) {
  p <- deseq_qc_ma(deseq_dir, outdir = deseq_dir, shrink_method = i) +
    ggtitle(glue::glue("shrink: {i}"))
  deseq_qc_add_sig_label(p)
})
patchwork::wrap_plots(p_list, ncol = 2, guides = "collect")
```

### 4.5 Volcano plot

```{r, fig.width = 7, fig.height = 7}
p_list <- lapply(names(res_lfc), function(i) {
  p <- deseq_qc_volcano(deseq_dir, outdir = deseq_dir, shrink_method = i) +
    ggtitle(glue::glue("shrink: {i}"))
  deseq_qc_add_sig_label(p)
})
patchwork::wrap_plots(p_list, ncol = 2, guides = "collect")
```

### 4.6 Scatter plot

```{r, fig.width = 8, fig.height = 7}
p_list <- lapply(names(res_lfc), function(i) {
  p <- deseq_qc_scatter(deseq_dir, outdir = deseq_dir, shrink_method = i) +
    ggtitle(glue::glue("shrink: {i}"))
  deseq_qc_add_sig_label(p)
})
patchwork::wrap_plots(p_list, ncol = 2, guides = "collect")
```


## 5 Detail report

```{r, report_list}
data_dir <- "data"
dir.create(data_dir, showWarnings = FALSE)
rpt_html <- list_hiseq_file(input_dir, "report_html", hiseq_type = TRUE)
rpt_html <- purrr::keep(rpt_html, file.exists)

if(length(rpt_html) > 0) {
  local_html <- sapply(rpt_html, function(f) {
    f_prefix <- basename(dirname(dirname(f)))
    f_html   <- file.path(data_dir, paste0(f_prefix, ".rnaseq_report.html"))
    if(! file.exists(f_html)) {
      file.copy(f, f_html)
    }
    f_html
  })
} else {
  local_html = NULL
}
# show in table
if(length(local_html) > 0) {
  data.frame(id = basename(dirname(dirname(rpt_html))),
             url = local_html
  ) %>%
    dplyr::mutate(link = paste0("<a href='", url, "' target='_blank'>", id, "</a>")) %>%
    dplyr::select(link) %>% 
    tibble::remove_rownames() %>%
    to_DT(mode = 3)
}
```


**END**
