---
title: "HiSeq report"
author: "Ming Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float:
      collapsed: no
    keep_md: false
  word_document:
    toc: yes
  pdf_document:
    toc: yes
params:
  input_dir: ""
---

```{r setup, include=FALSE}
suppressPackageStartupMessages(library(knitr))

knitr::opts_chunk$set(fig.width  = 12, 
                      fig.height = 8, 
                      fig.path   = "Figures/",
                      eval       = TRUE,
                      echo       = FALSE,
                      cache      = FALSE,
                      prompt     = FALSE,
                      tidy       = FALSE,
                      comment    = NA,
                      message    = FALSE,
                      warning    = FALSE,
                      eval       = T,
                      rownames.print = FALSE)
```


```{r load pkg}
library(hiseqr)
library(knitr)
library(ggplot2)
library(dplyr)
library(glue)
library(DT)
```


```{r init project, eval = T}
input_dir <- params$input_dir
# input_dir <- ".."
input_dir <- normalizePath(input_dir)
if(! is_hiseq_dir(input_dir, "rnaseq_rn")) {
  stop(paste0("Not a RNAseq single directory, _rn: ", input_dir))
}
```

## Config data

```{r config, eval = T, results = "asis"}
is_paired <- all(list_hiseq_file(input_dir, "is_paired", "r1"))
n_samples <- length(list_hiseq_file(input_dir, "rep_list", "_rn"))
genome    <- list_hiseq_file(input_dir, "genome", "_rn")
organism  <- genome
gtf       <- list_hiseq_file(input_dir, "gene_gtf", "_rn")
smp_name  <- list_hiseq_file(input_dir, "smp_name", "_rn")
is_mut    <- list_hiseq_file(input_dir, "is_mut", "_rn")
project_name <- list_hiseq_file(input_dir, "project_name", "_rn")
aligner   <- list_hiseq_file(input_dir, "aligner")
is_star   <- aligner %in% c("STAR", "hisat2")

msg <- glue::glue(
  "   smp_name: {smp_name}; ",
  "     genome: {genome}; ",
  "     is_mut: {is_mut}; ",
  " paired-end: {is_paired}",
  "project_dir: {project_name}",
  .sep = "\n"
)
print(msg)
```


```{r plot_data, eval = is_star}
df_trim  <- read_hiseq_trim_stat(input_dir)
df_align <- read_hiseq_align_stat(input_dir)
```


```{r loading data, eval = is_star}
pd <- hiseq_summary(input_dir, TRUE)

# summary
n_smp      <- length(list_hiseq_file(input_dir, "rep_list", "auto"))
if(inherits(pd$align, "data.frame") & inherits(pd$align, "data.frame")) {
  total      <- df_align %>% pull(total) %>% sum()
  map        <- df_align %>% pull(map) %>% sum()
  map_pct    <- round(map / total * 100, 2) %>% sum()
  unique_m   <- df_align %>% pull(unique) %>% sum()
  unique_pct <- round(unique_m / total * 100, 1)
  total      <- round(total / 1e6, 2)
  map        <- round(map / 1e6, 2)
  unique_m   <- round(unique_m / 1e6, 1) # Million
} else {
  total <- map <- unique_m <- 1
  unique_pct <- map_pct <- 100
}
```


## 1 Summary

```{r summary, results="asis", eval = is_star}
sum_text <- glue::glue("A total of {n_smp} samples; {total}M sequencing depth; ",
    "with {map}M ({map_pct}%) reads mapped to reference genome, including ",
    "{unique_m}M ({unique_pct}%) unique mapped.")

print(sum_text)
```


## 2 Results


### 2.1 Table 1. Trimming reads

```{r, eval = is_star}
pd$trim %>%
  dplyr::select(name, input, output, out_pct, rm_pct) %>%
  dplyr::rename(clean = output, 
                clean_pct = out_pct,
                short_pct = rm_pct) %>%
  hiseqr::to_DT()
```

```{r figure1_trim_reads, fig.height=3, fig.width=6, eval = is_star}
if(inherits(pd$trim, "data.frame")) {
  plot_hiseq_trim(pd$trim)
}
```


### 2.2 Table 2. Mapping reads

```{r, eval = is_star}
if(inherits(pd$align, "data.frame")) {
  pd$align %>%
    dplyr::select(name, total, map, unmap) %>%
    dplyr::mutate(map_pct = round(map / total * 100, 2)) %>%
    hiseqr::to_DT()
}
```

```{r figure1_align_reads, fig.height=3, fig.width=6, eval = is_star}
if(inherits(pd$align, "data.frame")) {
  plot_hiseq_align(pd$align, columns = c("map", "rRNA", "unmap"))
}
```

+ unique and multiple mapping reads

```{r, eval = is_star}
if(inherits(pd$align, "data.frame")) {
pd$align %>%
  dplyr::select(name, total, unique, multi, unmap) %>%
  dplyr::mutate(unique_pct = round(unique / total * 100, 2)) %>%
  hiseqr::to_DT()
}
```

```{r unique_multi_reads, fig.height=3, fig.width=6, eval = is_star}
if(inherits(pd$align, "data.frame")) {
plot_hiseq_align(pd$align, columns = c("unique", "multi", "unmap"))
}
```

### 2.3 RNAseq library type

Check the RNAseq library type, check if the reads from forward strand or reverse strand of mRNA;

Using `featureCounts` to count reads on mRNAs.

For strand-specific library

| sens | anti | library | example | 
| -- | -- | -- | -- |
| -s 1 | -s 2 | `read1: ++ -- / read2: -- ++` | - | 
| -s 2 | -s 1 | `read1: +- -+ / read2: +- -+` | dUTP library, NSR library |


```{r lib type, results="asis", eval = is_star}
lib1 = "read1: ++ -- / read2: -- ++"
lib2 = "read1: +- -+ / read2: +- -+"
lib3 = "read1, read2, non-strand-specific"

f <- list_hiseq_file(input_dir, "strandness_json", "_rn")
s <- jsonlite::read_json(f)
a = switch(as.character(s$sense),
        "0" = {print(lib3)},
        "1" = {print(lib1)},
        "2" = {print(lib2)})
# print(a)
```

### 2.4 Quantification

```{r, eval = is_star}
log2 <- glue::glue("Count reads on features:
                   Genome: {genome}
                      GTF: {basename(gtf)}")
print(log2)
```
**na**, indicate the `Not Assigned reads`.

```{r table2_sens_summary, eval = is_star}
p4 <- plot_rnaseq_featureCounts(input_dir)
# show table
df4 <- p4$data %>%
  dplyr::select(sample, strand, count, Status, total) %>%
  tidyr::pivot_wider(names_from = strand, values_from = count) %>%
  dplyr::mutate(sens_pct = round(sense / total * 100, 1),
                anti_pct = round(anti / total * 100, 1),
                na_pct   = round(na / total * 100, 1))
hiseqr::to_DT(df4)
```


**na**, indicate the `Not Assigned reads`.
```{r count strand, fig.height=3, fig.width=6}
print(p4)
```


### 2.5 Correlation between replicates

#### 1 Distribution of reads on gene body

#### 2 Correlation - PCA

#### 3 Correlation - heatmap

```{r, eval = is_star}
g1 <- list_hiseq_file(input_dir, "genebody_enrich_png", "_rn")
g2 <- list_hiseq_file(input_dir, "bam_cor_pca_png", "_rn")
g3 <- list_hiseq_file(input_dir, "bam_cor_heatmap_png", "_rn")
g  <- purrr::keep(c(g1, g2, g3), file.exists)
if(file.exists(g)) {
  include_graphics(g)
}
```

**END**


