---
title: "Check P7 structure"
author: "Ming Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float:
      collapsed: no
    keep_md: false
  word_document:
    toc: yes
  pdf_document:
    toc: yes
params:
  input_dir: ""
---

```{r setup, include=FALSE}
suppressPackageStartupMessages(library(knitr))

knitr::opts_chunk$set(fig.width  = 12, 
                      fig.height = 8, 
                      fig.path   = "Figures/",
                      eval       = TRUE,
                      echo       = FALSE,
                      cache      = FALSE,
                      prompt     = FALSE,
                      tidy       = FALSE,
                      comment    = NA,
                      message    = FALSE,
                      warning    = FALSE,
                      eval       = T,
                      rownames.print = FALSE)
```


```{r load pkg}
library(hiseqr)
library(knitr)
library(ggplot2)
library(fishualize)
library(patchwork)
library(dplyr)
library(glue)
```


```{r config, eval = T}
# main
input_dir <- params$input_dir
input_dir <- normalizePath(input_dir)
f_list    <- list.files(input_dir, "*stat.log$", full.names = TRUE)
if(length(f_list) == 0) {
  stop("No *stat.log files detected")
}
p_list    <- hiseqr::hiseq_p7(f_list)
```


## 1. HiSeq library structure 

```{r hiseq_lib_structure, fig.height=1, fig.width=8}
ft <- c("P5", "Insert", "barcode", "P7_a", "i7", "P7_b")
ds <- data.frame(
  hiseq = "TruSeq",
  score = c(3, 2, 1, 2, 1, 2),
  name  = ft
) %>%
  dplyr::mutate(name = forcats::fct_relevel(name, rev(ft)),
                label_x = cumsum(score) - 0.5 * score)

ds %>%
  ggplot(aes(score, hiseq, fill = name, group = name)) + 
  geom_col() + 
  scale_fill_manual(
    values = c("blue4", "red4", "blue4", "red4", "grey40", "blue4")) + 
  geom_text(aes(x = label_x, label = name), colour = "white", size =5) +
  ggtitle("Structure of TruSeq/Nextera library") + 
  theme_minimal() + 
  theme(legend.position = "none",
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank())
```

```
## TruSeq 122nt+insert, 129nt+insert(inline barcode)
P5:   AATGATACGGCGACCACCGAGATCTACACTCTTTCCCTACACGACGCTCTTCCGATCT, 58nt
barcode: 6nt + A
P7_a: AGATCGGAAGAGCACACGTCTGAACTCCAGTCAC, 34nt [TruSeq]
i7:   6nt  
P7_b: ATCTCGTATGCCGTCTTCTGCTTG, 24nt

## Nextera 128nt+insert
P5:   AATGATACGGCGACCACCGAGATCTACACTCGTCGGCAGCGTCAGATGTGTATAAGAGACAG, 62nt 
P7_a: CTGTCTCTTATACACATCTCCGAGCCCACGAGAC, 34nt  
i7:   8nt
P7_b: ATCTCGTATGCCGTCTTCTGCTTG, 24nt

```

## 1 The distribution of P7 components

### 1.1 Table1. P7 components

```{r}
df1 <- lapply(p_list, function(p) {
  p$p7_table
}) %>%
  dplyr::bind_rows() %>%
  dplyr::select(sample, total, group, pct) %>%
  dplyr::mutate(pct = paste0(pct, "%")) %>%
  tidyr::pivot_wider(names_from = "group", values_from = "pct")
df1 %>% hiseqr::to_DT()
```


### 1.2 Plots of P7 components

Check the frequency of each `i7` sequences; `x-axis` is percentage of sequence.

```{r p7_component, fig.width=7, fig.height=1.3}
p7_list <- lapply(p_list, function(p) { print(p$p7) })
```


## 2 The distribution of index7 sequence

### 2.1 Table2 the distribution of index7 sequence 

```{r}
df3 <- lapply(p_list, function(p) {
  p$i7_table
}) %>%
  dplyr::bind_rows() %>%
  dplyr::ungroup() %>%
  dplyr::select(rank, sample, total, demx_name, demx_pct) %>%
  dplyr::mutate(demx_label = paste0(demx_name, " (", demx_pct, "%)")) %>%
  dplyr::select(sample, total, demx_label, rank) %>%
  tidyr::pivot_wider(names_from = "rank", values_from = "demx_label")
  
df3 %>% hiseqr::to_DT()
```

### 2.2 Plots for index7 sequence

Check the frequency of each `barcode` sequences; `x-axis` is percentage of sequence.

```{r i7_sequence, fig.width=7, fig.height=1.5}
i7_list <- lapply(p_list, function(p) { print(p$i7) })
```


## 3 The distribution of barcode sequence

### 3.1 Table3 the distribution of barcode sequence

```{r}
df2 <- lapply(p_list, function(p) {
  p$bc_table
}) %>%
  dplyr::bind_rows() %>%
  dplyr::ungroup() %>%
  dplyr::select(rank, sample, total, demx_name, demx_pct) %>%
  dplyr::mutate(demx_label = paste0(demx_name, " (", demx_pct, "%)")) %>%
  dplyr::select(sample, total, demx_label, rank) %>%
  tidyr::pivot_wider(names_from = "rank", values_from = "demx_label")
  
df2 %>% hiseqr::to_DT()
```

### 3.2 Plots for barcode sequence 

```{r barcode_sequence, fig.width=7, fig.height=1.5}
bc_list <- lapply(p_list, function(p) { print(p$bc) })
```



>EOF









































































