---
title: "DESeq pair - compare"
author: "Ming Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float:
      collapsed: no
    keep_md: true
  word_document:
    toc: yes
  pdf_document:
    toc: yes
params:
  dirA: ""
  dirB: ""
---

```{r setup, include=FALSE}
suppressPackageStartupMessages(library(knitr))

knitr::opts_chunk$set(fig.width  = 12, 
                      fig.height = 8, 
                      fig.path   = "Figures/",
                      eval       = TRUE,
                      echo       = FALSE,
                      cache      = FALSE,
                      prompt     = FALSE,
                      tidy       = FALSE,
                      comment    = NA,
                      message    = FALSE,
                      warning    = FALSE,
                      eval       = T,
                      rownames.print = FALSE)
```


```{r load pkg}
library(hiseqr)
library(knitr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(glue)
library(DT)
library(ggvenn)
library(RColorBrewer)
library(ggalluvial)
library(patchwork)
```


```{r config}
# setwd("/data/yulab/wangming/work/yu_2020/projects/20200629_lxh_RNAseq/results/20200702_rnaseq_cmp/aaa")
# feature <- "gene"
# dirA    <- "../data/rnaseq_0310/shWhite_6h.vs.shAub-shAgo3_6h/"
# dirB    <- "../data/rnaseq_0310/shWhite_6h.vs.shPiwi-1_6h/"
dirA <- params$dirA
dirB <- params$dirB

## check
if(hiseqr::is_deseq_single_dir(dirA) & hiseqr::is_deseq_single_dir(dirB)) {
  message("Compare deseq_dir")
} else {
  stop("deseq_singl expected, failed")
}
```


```{r}
pd <- hiseqr::read_deseq_pair(dirA, dirB, "gene")
# save to file
rds_gene <- "deseq_pair.gene.rds"
saveRDS(pd, rds_gene)
```

## Note

```{r}
msg <- glue::glue("Sample info:
                  GroupA: {basename(pd$dirA)}
                  GroupB: {basename(pd$dirB)}")
print(msg)
```

## 1 Gene

## 1.1 Number of significant changed genes

```{r}
ps <- hiseqr::deseq_pair_stat(.pd_rds = rds_gene, feature = "gene")
hiseqr::get_DT_table(ps$sig_count_table, 1)
```

## 1.2 Overlap between groups

```{r, fig.width = 10, fig.height = 4}
po <- hiseqr::deseq_pair_overlap_plot(.pd_rds = rds_gene) # 
po$merge
```

## 1.3 Overlap between sig genes

```{r, fig.width = 10, fig.height = 6}
po$up_down
```

## 1.4 Heatmap - sig genes

```{r, fig.width = 5, fig.height = 8}
hiseqr::deseq_pair_heatmap(.pd_rds = rds_gene)
```

## 1.5 Alluvial - sig genes

```{r, fig.width = 6, fig.height = 4}
hiseqr::deseq_pair_alluvial_plot(.pd_rds = rds_gene)
```

## 1.6 Sig gene patterns

```{r}
df2 <- ps$sig_list %>%
  dplyr::group_by(trend) %>%
  dplyr::summarise(count = n())

hiseqr::get_DT_table(df2, 1)
```

## 1.7 Sig gene list

```{r}
df3 <- ps$sig_list %>%
  dplyr::mutate(symbol = hiseqr::gene_to_symbol(Gene, "dm3", "SYMBOL"))

hiseqr::get_DT_table(df3, 1, pageLength = 20)
```


## 2 TE

## 2.1 Number of significant changed TEs

```{r}
pd     <- hiseqr::read_deseq_pair(dirA, dirB, "te")
rds_te <- "deseq_pair.te.rds"
saveRDS(pd, rds_te)
ps     <- hiseqr::deseq_pair_stat(.pd_rds = rds_te)
hiseqr::get_DT_table(ps$sig_count_table, 1)
```

## 2.2 Compare TE expression

GroupB_exp was normalized by GroupA_ctl.

```{r, fig.width = 5, fig.height = 10}
hiseqr::deseq_pair_heatmap_te(dirA, dirB)
```

## 2.3 Overlap between groups

```{r, fig.width = 10, fig.height = 4}
po <- hiseqr::deseq_pair_overlap_plot(.pd_rds = rds_te) # 
po$merge
```

## 2.4 Overlap between sig TEs

```{r, fig.width = 10, fig.height = 6}
po$up_down
```

## 2.5 Heatmap - sig TEs

```{r, fig.width = 5, fig.height = 8}
hiseqr::deseq_pair_heatmap(.pd_rds = rds_te)
```

## 2.6 Alluvial - sig TEs

```{r, fig.width = 6, fig.height = 4}
hiseqr::deseq_pair_alluvial_plot(.pd_rds = rds_te)
```

## 2.7 Sig gene patterns

```{r}
df2 <- ps$sig_list %>%
  dplyr::group_by(trend) %>%
  dplyr::summarise(count = n())

hiseqr::get_DT_table(df2, 1)
```

## 2.8 Sig gene list

```{r}
df3 <- ps$sig_list 
hiseqr::get_DT_table(df3, 1, pageLength = 20)
```


## 3 piRNA cluster

## 3.1 Number of significant changed piRNA cluster

```{r}
pd     <- hiseqr::read_deseq_pair(dirA, dirB, "piRNA_cluster")
rds_pi <- "deseq_pair.te.rds"
saveRDS(pd, rds_pi)
ps     <- hiseqr::deseq_pair_stat(.pd_rds = rds_pi)
hiseqr::get_DT_table(ps$sig_count_table, 1)
```

## 3.2 Overlap between groups

```{r, fig.width = 10, fig.height = 4}
po <- hiseqr::deseq_pair_overlap_plot(.pd_rds = rds_pi) # 
po$merge
```

## 3.3 Overlap between sig piRNA cluster

```{r, fig.width = 10, fig.height = 6}
po$up_down
```

## 3.4 Heatmap - sig piRNA cluster

```{r, fig.width = 5, fig.height = 8}
hiseqr::deseq_pair_heatmap(.pd_rds = rds_pi)
```

## 3.5 Alluvial - sig piRNA cluster

```{r, fig.width = 6, fig.height = 4}
hiseqr::deseq_pair_alluvial_plot(.pd_rds = rds_pi)
```

## 3.6 Sig gene patterns

```{r}
df2 <- ps$sig_list %>%
  dplyr::group_by(trend) %>%
  dplyr::summarise(count = n())

hiseqr::get_DT_table(df2, 1)
```

## 3.7 Sig gene list

```{r}
df3 <- ps$sig_list 
hiseqr::get_DT_table(df3, 1, pageLength = 20)
```




**END**
